---
title: "Collecting Data: Web, APIs, Open Sources"
subtitle: "E1493 - Data Journalism"
author: "Simon Munzert"
institution: "Hertie School"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: false
    theme: cosmo
    code-fold: false
    code-tools: true
    df-print: paged
    fig-width: 8
    fig-height: 5
    highlight-style: github
    embed-resources: true
    link-external-newwindow: true
execute:
  eval: false
  warning: false
  message: false
---

```{r}
#| eval: true
#| include: false
#| label: load-packages
  
pkgs <- c("httr2", "jsonlite", "rvest", "tidyverse", "selenider")
lapply(pkgs, library, character.only = TRUE)
```

::: {.callout-note}
## Before you start

Install the packages used in this workshop:

```r
install.packages(c("httr2", "jsonlite", "rvest", "tidyverse", "selenider"))

# selenider needs a browser driver — run once:
# For Chrome (default):
selenider::install_selenider()           # installs chromedriver via wdman
# For Firefox:
selenider::install_selenider("firefox")  # installs geckodriver via wdman
```

You will also need **Google Chrome** or **Firefox** installed. Chrome is the default in this workshop; swap `browser = "firefox"` in the session setup if you prefer Firefox.
:::

---

# Part 1: HTTP and httr2 Refresher {#http-refresher}

Every time you load a webpage or call an API, your computer and a server are exchanging **HTTP messages**. Understanding the basics of those messages is the foundation for everything that follows.

## The request–response cycle

An HTTP interaction always follows the same pattern:

1. Your code sends a **request** to a URL
2. The server processes it and sends back a **response**
3. The response has a **status code**, **headers**, and a **body**

```
Client                          Server
  |                               |
  |  GET /w/rest.php/v1/search    |
  |  Host: en.wikipedia.org       |
  |  Accept: application/json     |
  | ----------------------------→ |
  |                               |
  |  200 OK                       |
  |  Content-Type: application/json
  |  {"pages": [...]}             |
  | ←---------------------------- |
```

## HTTP methods

The two methods you'll encounter most often:

**GET** — retrieve data. Parameters go in the URL as query strings:
```
https://api.genderize.io?name=Anna

https://www.google.com/search?q=Berlin
```

**POST** — send data to the server. Parameters go in the **request body** (usually JSON). Used when the query is complex or when the server is doing computation:
```
POST https://api.ted.europa.eu/v3/notices/search
Body: {"query": "TD=7 AND buyer-country=DEU", "limit": 25}
```

## Status codes

The three-digit code tells you immediately whether something worked:

| Code | Meaning | What it might imply |
|---|---|---|
| `200` | OK | Everything worked! |
| `400` | Bad Request | Your request is malformed |
| `401` | Unauthorized | Maybe need an API key |
| `403` | Forbidden | Access denied |
| `404` | Not Found | Wrong URL? |
| `429` | Too Many Requests | Slow down — you're being rate-limited |
| `500` | Server Error | Problem on their end |

## Headers

Headers are key-value metadata attached to both requests and responses. Common ones you'll set or read:

**Request headers** (you send):

- `Accept: application/json` — tell the server you want JSON back
- `Accept-Encoding: gzip, deflate` — tell the server which compression formats R can handle (important: omit `br`/brotli, which libcurl does not support by default)
- `User-Agent` — identifies your client; good practice to set this
- `Authorization: Bearer <token>` — for authenticated APIs

**Response headers** (server sends back):

- `Content-Type` — format of the response body
- `X-Rate-Limit-Remaining` — how many requests you have left
- `X-RateLimit-Reset` — when the limit resets

## httr2 in practice

`httr2` is the modern R package for HTTP. Its key design is a **pipeline**: build a request object, add modifiers, then perform it.

```{r}
#| label: httr2-intro
library(httr2)

# Build → modify → perform → extract
resp <- request("https://api.genderize.io") |>  # base URL
  req_url_query(name = "Anna") |>               # add ?name=Anna
  req_headers(Accept = "application/json") |>   # set a header
  req_timeout(10) |>                            # fail after 10s
  req_perform()                                 # fire the request

# Inspect the response
resp_status(resp)       # 200
resp_content_type(resp) # "application/json"
resp_headers(resp)      # all response headers

# Extract the body
resp_body_string(resp)  # raw JSON string
resp_body_json(resp)    # parsed into R list via jsonlite
```

### Key httr2 modifiers

| Function | What it does |
|---|---|
| `req_url_query(key = val)` | Append `?key=val` to the URL |
| `req_headers(Key = "val")` | Set request headers |
| `req_body_json(list(...))` | Send a JSON body (POST) |
| `req_auth_bearer_token(tok)` | Add `Authorization: Bearer` header |
| `req_throttle(n / 60)` | Limit to n requests per minute |
| `req_retry(max_tries = 3)` | Auto-retry on failure |
| `req_timeout(seconds)` | Set a timeout |

### Reading response headers

```{r}
#| label: read-headers
# Single header
resp |> resp_header("Content-Type")

# All headers as a named list
resp |> resp_headers()
```

::: {.callout-tip}
## Where does the Network tab fit in?

When you use your browser to interact with a website, every HTTP request your browser makes is visible in the **Network tab** of the developer tools. This means you can inspect the exact URL, method, headers, and body of any request — and then reproduce it in R. That is the entire basis of the undocumented API technique in Part 3.
:::

## Exercises {#exercises-p1}

**Exercise 1 (easy) — Inspect a response with `httpbin`**

[httpbin.org](https://httpbin.org) is a free testing service that echoes back whatever request you send it. Send a GET request to `https://httpbin.org/get` with two query parameters of your choice and inspect the result. What does `$args` contain? What user-agent does R report?

::: {.callout-note collapse="true"}
## Solution

```r
resp <- request("https://httpbin.org/get") |>
  req_url_query(city = "Berlin", year = 2025) |>
  req_perform()

resp_status(resp)
# [1] 200

body <- resp_body_json(resp)
body$args
# $city  "Berlin"
# $year  "2025"

body$headers$`User-Agent`
# "libcurl/8.x httr2/1.x r-curl/5.x"
```

The `$args` field echoes your query parameters back — confirming the server received them. The `User-Agent` shows R's default identifier; override it with `req_headers("User-Agent" = "MyScript/1.0")` for polite API usage.
:::

---

**Exercise 2 (intermediate) — Send a POST request with a JSON body**

Some APIs use POST rather than GET, with data sent in the request body instead of the URL. Use `req_body_json()` to POST a small payload to `https://httpbin.org/post`. Which field in the response contains your data? What `Content-Type` did httr2 set automatically?

::: {.callout-note collapse="true"}
## Solution

```r
resp <- request("https://httpbin.org/post") |>
  req_body_json(list(name = "Anna", country = "DE", year = 2025)) |>
  req_perform()

body <- resp_body_json(resp)

body$json
# $name    "Anna"
# $country "DE"
# $year    2025

body$headers$`Content-Type`
# "application/json"
```

`req_body_json()` serialises the R list to JSON and sets `Content-Type: application/json` automatically — no manual header needed. Contrast with `req_url_query()`: GET puts parameters in the URL; POST puts them in the body. POST is necessary when the payload is large or sensitive, or when the server is doing computation rather than simple retrieval.
:::

---

# Part 2: API Refresher {#refresher}

Before we go hunting for hidden APIs, let's make sure the basics are solid. An **API** (Application Programming Interface) is simply a structured way for two programs to talk to each other. On the web, that usually means: your code sends an HTTP request to a URL, and the server responds with data — nowadays almost always in **JSON** format.

## JSON in R

JSON is the lingua franca of web APIs. The standard R package for parsing it is **jsonlite**, which has been the community default for a decade. It's robust, well-maintained, and handles the tricky R-specific cases (data frames, NAs, nested lists) gracefully.

```{r}
#| eval: true

library(jsonlite)

# Parse a JSON string
fromJSON('{"name": "Anna", "gender": "female", "probability": 0.97}')

# Convert R objects to JSON
toJSON(list(name = "Anna", country = "DE"), auto_unbox = TRUE)
```

## Example: the Genderize.io API

[Genderize.io](https://genderize.io) predicts the likely gender of a name, along with a confidence score (`probability`) and the number of data points in its training database (`count`). It requires **no API key** for up to 100 names/day.

It's also a useful object for thinking about **algorithmic fairness**: the model is trained heavily on Western names, so confidence is high for "James" or "Sophie" but degrades sharply for names from South Asia, East Africa, or East Asia.

Before interacting with an API, it's worth checking out its documentation to understand the endpoints, parameters, and response structure. For Genderize.io, the docs are at [https://genderize.io/documentation](https://genderize.io/documentation).

### Single name lookup

```{r}
library(httr2)
library(jsonlite)

resp <- request("https://api.genderize.io") |>
  req_url_query(name = "Anna") |>
  req_perform()

resp |>
  resp_body_string() |>
  fromJSON()
# $count    $name    $gender    $probability
# 1673584   "Anna"   "female"   0.9          
```

### Checking your rate limit

The API returns rate-limit information in the **response headers**. With 100 free requests/day, it's worth checking before running a big batch:

```{r}
# Inspect rate-limit headers from the last response
list(
  limit     = resp |> resp_header("X-Rate-Limit-Limit"),
  remaining = resp |> resp_header("X-Rate-Limit-Remaining"),
  reset_in  = resp |> resp_header("X-Rate-Limit-Reset") |> as.integer() |> (\(s) paste(s, "seconds"))()
)
```

### Batch requests and a helper function

The API accepts up to 10 names per request using repeated `name[]` parameters. Before writing our own wrapper, it's worth checking whether a dedicated R client already exists — packages like `genderizeR` were built exactly for this. However, `genderizeR` is no longer under active development and currently causes installation issues due to a corrupted dependency (`stringi`) on some systems. This is a good general lesson: **always look for existing wrappers and clients, but check when they were last updated and whether issues are being addressed before relying on them.**

Here's a small wrapper built directly on `httr2` that handles batching and returns a tidy data frame:

```{r}
#| eval: true
  
genderize <- function(names) {
  batches <- split(names, ceiling(seq_along(names) / 10))

  map(batches, \(batch) {
    resp <- request("https://api.genderize.io") |>
      req_url_query(!!!setNames(as.list(batch), rep("name[]", length(batch)))) |>
      req_throttle(10 / 60) |>
      req_perform()

    resp |>
      resp_body_string() |>
      fromJSON()
  }) |>
    list_rbind()
}

# Try it
genderize(c("James", "Sophie", "Amara")) |>
  select(name, gender, probability, count)
```

### Mini-audit: confidence by name origin

```{r}
#| eval: true

names_df <- tibble(
  name   = c("James", "Sophie", "Ingrid",
             "Amara", "Chioma", "Kwame",
             "Wei",   "Yuki",   "Ji-ho"),
  region = c(rep("W. European", 3),
             rep("Sub-Saharan African", 3),
             rep("East Asian", 3))
)

results <- genderize(names_df$name) |>
  select(name, gender, probability, count) |>
  left_join(names_df, by = "name")

ggplot(results, aes(x = reorder(name, probability), y = probability, fill = region)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Genderize.io: Prediction confidence by name origin",
    x = NULL, y = "Probability", fill = "Region"
  ) +
  theme_minimal()
```

::: {.callout-tip}
## Discussion

- What do low `count` values tell you about the training data?
- If a company uses genderize.io to personalise emails, what's the downstream effect of low-confidence predictions?
- How would you design a more rigorous audit — what would your name sample look like, and what would you measure?
:::

## Exercises {#exercises-p2}

**Exercise 1 (easy) — Look up your own name**

Query genderize.io with your own first name (or a colleague's). Then query it again with a name from a very different cultural background — e.g. "Kwabena", "Priyanka", or "Seun". Compare the `probability` and `count` values. What do the differences tell you?

::: {.callout-note collapse="true"}
## Solution

```r
genderize(c("Simon", "Kwabena")) |>
  select(name, gender, probability, count)

#   name    gender probability  count
#   Simon   male   0.99         304891
#   Kwabena male   0.97         153
```

The `count` for "Kwabena" (a common Akan name from Ghana) is orders of magnitude lower than for "Simon". This low count drives the model's uncertainty in edge cases and reflects the Western skew of the training data — even when the `probability` happens to look high, the model has seen very few examples of that name.
:::

---

**Exercise 2 (intermediate) — Audit German party leaders**

The Bundestag has six parliamentary groups (Fraktionen). Feed the first names of their current leaders into `genderize()` and visualise the results. What does the confidence pattern look like — and why might it be misleading to use this API for political reporting on non-Western politicians?

::: {.callout-note collapse="true"}
## Solution

```r
leaders <- tibble(
  name  = c("Friedrich", "Olaf", "Robert", "Alice", "Sahra", "Anton"),
  party = c("CDU/CSU", "SPD", "Grüne", "AfD", "BSW", "FDP")
)

genderize(leaders$name) |>
  left_join(leaders, by = "name") |>
  select(name, party, gender, probability, count) |>
  arrange(probability)
```

All six names are common in German and score very high confidence (>0.97) — which is expected. The exercise is more revealing when you swap in names of politicians from minority backgrounds or from non-European countries: confidence and `count` drop sharply, making the model unreliable for exactly the groups it's most likely to misclassify. This is the core fairness concern with name-based gender inference in journalism.
:::

---

# Part 3: Undocumented APIs {#part3}

> *"Most APIs are undocumented and hidden in plain sight."*  
> — Leon Yin, [Inspect Element](https://inspectelement.org/apis.html)

When you interact with a modern website — filtering search results, browsing an encyclopedia, or querying a government portal — your browser is quietly making structured data requests behind the scenes. These are **API calls**, and the data they return is often richer, cleaner, and more machine-readable than anything you could scrape from HTML.

The trick is finding them.

## What is an undocumented API?

A documented API comes with a manual. The EU Open Data Portal, the U.S. Census Bureau, and the World Bank all publish official APIs with authentication keys, rate limits, and endpoint documentation.

Undocumented APIs are the ones developers built for their own front-end — they power the search boxes, the dropdown filters, and the dynamic tables you see on websites. No one advertised them. But they're there, and they speak JSON.

Some investigations are **only possible** once you find one of these hidden endpoints.

---

## Case study: Wikipedia search

[Wikipedia](https://en.wikipedia.org) runs on **MediaWiki**, and behind its search box lies a clean, well-structured REST API. When you type into the Wikipedia search field and autocomplete suggestions appear, your browser is calling an internal endpoint — one that is not prominently advertised but is entirely open to use.

This makes Wikipedia an ideal first target: the API is stable, requires no authentication, and returns well-structured JSON. The same technique you learn here applies directly to more complex targets — government databases, corporate transparency portals, parliamentary records.

---

## Tutorial: Finding and using undocumented APIs {#api-tutorial}

We'll use Wikipedia's search as our target. The goal is to intercept the API call the website makes when you search, and reproduce it in R.

### Step 1: Open the browser developer tools {#p3-step1}

Navigate to [en.wikipedia.org](https://en.wikipedia.org/wiki/Main_Page) and open the developer tools:

- **Mac**: `Cmd + Option + I`
- **Windows/Linux**: `F12` or `Ctrl + Shift + I`

Click the **Network** tab.

### Step 2: Trigger a search request {#p3-step2}

Click **Fetch/XHR** in the filter bar to show only API calls. Then type **"Hertie"** into the Wikipedia search box at the top of the page. New entries will appear in the Network panel as autocomplete results load.

### Step 3: Find the data request {#p3-step3}

Look for a request with a URL containing `/rest.php/v1/search/`. Click it and inspect:

- **Headers tab** → method is `GET`, URL is `https://en.wikipedia.org/w/rest.php/v1/search/title?q=Hertie&limit=10`
- **Response tab** → a JSON object with a `pages` array, each entry containing `id`, `title`, `description`, and `thumbnail`

### Step 4: Copy as cURL and convert {#p3-step4}

Right-click the request → **Copy** → **Copy as cURL**.

Paste into [curlconverter.com](https://curlconverter.com/) and select **R**. You'll get something like this:

```{r}
#| label: curlconverter-raw
library(httr2)

wiki_out <-
  request("https://en.wikipedia.org/w/rest.php/v1/search/title") |>
  req_url_query(
    q     = "Hertie",
    limit = "10"
  ) |>
  req_headers(
    `User-Agent`      = "Mozilla/5.0 (Macintosh; Intel Mac OS X 10.15; rv:147.0) Gecko/20100101 Firefox/147.0",
    Accept            = "application/json",
    `Accept-Language` = "en-US,en;q=0.9",
    `Accept-Encoding` = "gzip, deflate",   # ← remove br/zstd: libcurl can't decode them
    Referer           = "https://en.wikipedia.org/wiki/Main_Page",
    `Sec-GPC`         = "1",
    Connection        = "keep-alive",
    # Cookie          = "..."  ← remove: cookies are session-specific and unnecessary
    `Sec-Fetch-Dest`  = "empty",
    `Sec-Fetch-Mode`  = "cors",
    `Sec-Fetch-Site`  = "same-origin",
    DNT               = "1"
  ) |>
  req_perform()
```

::: {.callout-tip}
## Cleaning up curlconverter output

curlconverter faithfully copies **everything** your browser sent — including session cookies, brotli encoding flags, and browser fingerprinting headers. Most of these are unnecessary for the API call to work. Two things always need cleaning:

1. **Remove or replace the Cookie header** — it is session-specific and will expire. Most public APIs don't need it.
2. **Change `Accept-Encoding` to `"gzip, deflate"`** — browsers advertise brotli (`br`) by default, but R's libcurl cannot decode it and will throw an error.

Everything else (Sec-Fetch-*, DNT, Priority) is browser metadata that servers typically ignore. You can remove it to keep your code clean.
:::

### Step 5: A minimal working request {#p3-step5}

Once you know what the API needs, strip it down. For Wikipedia's search endpoint, all that's required is the query and a sensible `User-Agent`:

```{r}
#| label: minimal-request
#| eval: true
library(httr2)
library(tidyverse)

wiki_out <-
  request("https://en.wikipedia.org/w/rest.php/v1/search/title") |>
  req_url_query(q = "Hertie", limit = 10) |>
  req_headers(
    `User-Agent`      = "DataJournalismCourse/1.0 (hertie-school.org)",
    Accept            = "application/json",
    `Accept-Encoding` = "gzip, deflate"
  ) |>
  req_perform()

resp_status(wiki_out)  # 200 = success
```

### Step 6: Parse the response {#p3-step6}

```{r}
#| label: parse-wiki
#| eval: true

wiki_df <-
  wiki_out |>
  resp_body_json(simplifyVector = TRUE) |>
  pluck("pages")

wiki_df
```

`simplifyVector = TRUE` passes a flag to jsonlite that collapses lists of uniform objects directly into a data frame. The `thumbnail` column comes out as a nested data frame; flatten it with `unnest_wider()`:

```{r}
#| label: flatten-wiki
#| eval: true
wiki_df |>
  unnest_wider(thumbnail, names_sep = "_") |>
  select(id, title, description, thumbnail_url, thumbnail_width, thumbnail_height)
```

### Step 7: Explore the API further {#p3-step7}

```{r}
#| label: wiki-explore

# Full-text search (searches article bodies, not just titles)
request("https://en.wikipedia.org/w/rest.php/v1/search/page") |>
  req_url_query(q = "Hertie School", limit = 5) |>
  req_headers(`User-Agent` = "DataJournalismCourse/1.0",
              `Accept-Encoding` = "gzip, deflate") |>
  req_perform() |>
  resp_body_json(simplifyVector = TRUE) |>
  pluck("pages") |>
  select(title, description, excerpt)
```

```{r}
#| label: wiki-paginate

# Paginate: use the `offset` parameter to retrieve the next page of results
page2 <-
  request("https://en.wikipedia.org/w/rest.php/v1/search/page") |>
  req_url_query(q = "Hertie School", limit = 5, offset = 5) |>
  req_headers(`User-Agent` = "DataJournalismCourse/1.0",
              `Accept-Encoding` = "gzip, deflate") |>
  req_perform() |>
  resp_body_json(simplifyVector = TRUE) |>
  pluck("pages")
```

::: {.callout-caution}
## Responsible use

- Wikipedia's API is free and open, but the [Wikimedia API etiquette](https://www.mediawiki.org/wiki/API:Etiquette) asks that you identify your client in the `User-Agent` header and keep requests reasonable.
- Use `req_throttle()` if you're making many calls in a loop.
- Cache results locally — don't re-fetch data you already have.
:::

## Exercises {#exercises-p3}

**Exercise 1 (easy) — Search for your city**

Use the Wikipedia title-search endpoint to retrieve the top 10 results for your hometown or a city of your choice. How many results have a non-empty `description` field?

::: {.callout-note collapse="true"}
## Solution

```r
city_results <-
  request("https://en.wikipedia.org/w/rest.php/v1/search/title") |>
  req_url_query(q = "Berlin", limit = 10) |>
  req_headers(`User-Agent`      = "DataJournalismCourse/1.0",
              `Accept-Encoding` = "gzip, deflate") |>
  req_perform() |>
  resp_body_json(simplifyVector = TRUE) |>
  pluck("pages")

city_results |>
  select(title, description) |>
  mutate(has_description = !is.na(description) & description != "")

# Count
sum(!is.na(city_results$description))
```

The title-search endpoint matches article titles, not body text — so "Berlin" returns the city article first, then districts, suburbs, and disambiguation pages. Articles for the main topic usually have descriptions; stubs and disambiguation pages often don't.
:::

---

**Exercise 2 (intermediate) — Compare title search vs. full-text search**

Run the same query ("data journalism") through both the title-search endpoint (`/v1/search/title`) and the full-text search endpoint (`/v1/search/page`). Compare the results. Do they overlap? Which endpoint returns more relevant articles for a research task?

::: {.callout-note collapse="true"}
## Solution

```r
make_wiki_request <- function(endpoint, query, limit = 10) {
  request(paste0("https://en.wikipedia.org/w/rest.php/v1/search/", endpoint)) |>
    req_url_query(q = query, limit = limit) |>
    req_headers(`User-Agent`      = "DataJournalismCourse/1.0",
                `Accept-Encoding` = "gzip, deflate") |>
    req_perform() |>
    resp_body_json(simplifyVector = TRUE) |>
    pluck("pages") |>
    select(title, description) |>
    mutate(endpoint = endpoint)
}

title_results <- make_wiki_request("title", "data journalism")
page_results  <- make_wiki_request("page",  "data journalism")

combined <- bind_rows(title_results, page_results)

# Articles in both sets
intersect(title_results$title, page_results$title)

# Articles only in full-text search
setdiff(page_results$title, title_results$title)
```

The title-search endpoint only matches articles whose *title* contains "data journalism" — a small set. The full-text endpoint searches article bodies and returns a broader, richer set of related articles (investigative journalism, computational journalism, newsroom tools). For exploratory research, full-text is usually more useful; for disambiguation lookups, title search is more precise.
:::

---

### ✋ Your turn {#your-turn-p3}

1. Go to [tiktok.com](https://www.tiktok.com/), open the Network tab (filter by **Fetch/XHR**), and type a search query into the search box. Look for requests that return JSON containing search suggestions or results.
2. Copy one of those requests as cURL, convert it at curlconverter.com, clean up the headers (remove cookies, fix Accept-Encoding), and execute in R.
3. Parse the response. What does the data structure look like — are you seeing autocomplete suggestions, full results, or something else?
4. Try modifying the query parameter to a different search term. Does the same endpoint still work, or do you need to adjust other parameters?
5. Write a short paragraph: what story could you tell by systematically collecting TikTok search suggestions for a set of politically or socially relevant queries?

---

# Part 4: Browser Automation {#part4}

> *"Browser automation is a fundamental web scraping technique for building your own dataset. It is essential for investigating personalization, working with rendered elements, and waiting for scripts and code to execute on a web page."*  
> — Piotr Sapiezynski & Leon Yin, [Inspect Element](https://inspectelement.org/browser_automation.html)

Some websites do not have a convenient hidden API. The content is rendered by JavaScript after the page loads, injected into the DOM by a front-end framework — React, Vue, Angular — that `rvest` and `read_html()` will never see.

For these sites, you need **browser automation**: R controlling an actual Chrome browser, clicking buttons and waiting for content to appear, exactly as a human would.

## When do you need browser automation?

Prefer the undocumented API approach (Part 3) whenever possible — it is faster, lighter, and more reliable. Use browser automation when:

1. **The content is rendered dynamically** and does not appear in `view-source:`.
2. **You need to simulate interaction** — logging in, clicking a "load more" button, filling a form, or accepting a cookie banner.
3. **The site detects and blocks simple HTTP requests** but allows real browsers.

## Case study: YouTube search result audits

YouTube is built almost entirely in React — virtually nothing you care about appears in the raw HTML source. Every video title, channel name, view count, and thumbnail URL is injected into the DOM by JavaScript after the page loads. This makes `rvest` alone useless; browser automation is the only practical approach. YouTube does offer an official [Data API](https://paulcbauer.github.io/apis_for_social_scientists_a_review/youtube-api.html), but it returns aggregated metadata rather than the ranked search results and recommendation sequences that matter for algorithmic audits — making browser automation the necessary choice for this kind of research.

Researchers studying algorithmic amplification have used browser automation to systematically collect YouTube search results for sets of politically or scientifically contested queries — comparing what the platform surfaces for terms like "vaccine safety", "climate change", or "migration" across different languages and regions. The resulting datasets have been used to audit compliance with the EU's Digital Services Act (DSA), which requires platforms to assess and mitigate risks of "systemic" content recommendation.

Mozilla Foundation's [YouTube Regrets](https://www.mozillafoundation.org/en/youtube/findings/) project and AlgorithmWatch's [Automating Society](https://algorithmwatch.org/en/automating-society/) reports are prominent examples of this research genre. Both rely on automated browsers to collect the evidence that underpins their policy recommendations.

---

## Tutorial: Browser automation with `selenider` {#automation-tutorial}

[`selenider`](https://ashbythorpe.github.io/selenider/) is a modern R package that wraps Chrome via ChromeDriver. It has a clean, pipe-friendly API and excellent error messages — much easier to use than `RSelenium`.

::: {.callout-note}
## What is ChromeDriver?

**ChromeDriver** is a small server that acts as a bridge between your R code and a real Chrome browser. When `selenider` opens a browser window, it is actually sending commands to ChromeDriver, which translates them into Chrome's internal automation protocol (the [Chrome DevTools Protocol](https://chromedevtools.github.io/devtools-protocol/)). ChromeDriver must match the version of Chrome installed on your machine — `selenider::install_selenider()` handles this automatically by downloading the correct driver via the `wdman` package.
:::

### Step 1: Start a browser session {#p4-step1}

```{r}
#| label: selenider-setup
library(tidyverse)
library(rvest)
library(selenider)

session <- selenider_session(
  "chromote",
  options = chromote_options(headless = FALSE, width = 1280, height = 900)
)
```

### Step 2: Navigate to YouTube and search {#p4-step2}

Try fetching a YouTube search page the ordinary way first, so you can see why browser automation is necessary:

```{r}
#| label: why-not-rvest

library(rvest)

page <- read_html("https://www.youtube.com/results?search_query=climate+change")
page |> html_elements("h3") |> html_text()
# → character(0)   ← nothing there
```

Now with browser automation:

```{r}
#| label: navigate

query <- "climate change"

open_url(
  paste0("https://www.youtube.com/results?search_query=",
         URLencode(query, reserved = TRUE))
)

Sys.sleep(3)

# Handle the cookie consent popup — click "Reject all" if it appears
reject_btn <- s('#content > div.body.style-scope.ytd-consent-bump-v2-lightbox > div.eom-buttons.style-scope.ytd-consent-bump-v2-lightbox > div:nth-child(1) > ytd-button-renderer:nth-child(1) > yt-button-shape > button > yt-touch-feedback-shape > div.yt-spec-touch-feedback-shape__fill')
if (is_present(reject_btn)) {
  elem_click(reject_btn)
  Sys.sleep(2)
}

Sys.sleep(2)
```

### Step 3: Find elements on the rendered page {#p4-step3}

```{r}
#| label: find-elements

result_cards <- ss("ytd-video-renderer")
length(result_cards)
```

::: {.callout-tip}
## Reading the page source after JavaScript runs

After navigation and any interaction, extract the *rendered* HTML source and hand it to `rvest`:

```r
html <- read_html(get_page_source(session))
```

This captures what is actually displayed in the browser — not the empty shell that `read_html(url)` would return directly.
:::

### Step 4: Extract data from rendered HTML {#p4-step4}

```{r}
#| label: extract-html

html <- get_page_source()

titles <- html |>
  html_elements("ytd-video-renderer #video-title") |>
  html_text(trim = TRUE)

urls <- html |>
  html_elements("ytd-video-renderer #video-title") |>
  html_attr("href") |>
  (\(x) paste0("https://www.youtube.com", x))()

# YouTube renders two yt-formatted-string elements per channel
# (visible + screen-reader copy); take every other one
channels_raw <- html |>
  html_elements("ytd-video-renderer ytd-channel-name yt-formatted-string") |>
  html_text(trim = TRUE)
channels <- channels_raw[seq(1, length(channels_raw), by = 2)]

metadata <- html |>
  html_elements("ytd-video-renderer #metadata-line span") |>
  html_text(trim = TRUE)

view_counts  <- metadata[seq(1, length(metadata), by = 2)]
upload_dates <- metadata[seq(2, length(metadata), by = 2)]

results <- tibble(
  title    = titles,
  channel  = channels,
  views    = view_counts,
  uploaded = upload_dates,
  url      = urls
)

results
```

### Step 5: Collect results at scale with scrolling {#p4-step5}

```{r}
#| label: scroll-and-compare

collect_search_results <- function(query, n_scrolls = 3) {
  open_url(
    paste0("https://www.youtube.com/results?search_query=",
           URLencode(query, reserved = TRUE))
  )
  Sys.sleep(3)

  reject_btn <- s('#content > div.body.style-scope.ytd-consent-bump-v2-lightbox > div.eom-buttons.style-scope.ytd-consent-bump-v2-lightbox > div:nth-child(1) > ytd-button-renderer:nth-child(1) > yt-button-shape > button > yt-touch-feedback-shape > div.yt-spec-touch-feedback-shape__fill')
  if (is_present(reject_btn)) {
    elem_click(reject_btn)
    Sys.sleep(2)
  }

  # Scroll to the last visible result card to trigger lazy-loading.
  # YouTube's infinite scroll fires when the last card enters the viewport,
  # so elem_scroll_to() is more reliable than raw JS window.scrollBy().
  for (i in seq_len(n_scrolls)) {
    cards <- ss("ytd-video-renderer")
    elem_scroll_to(cards[[length(cards)]])
    Sys.sleep(2)
  }

  html <- get_page_source()

  titles <- html |> html_elements("ytd-video-renderer #video-title") |> html_text(trim = TRUE)
  channels_raw <- html |> html_elements("ytd-video-renderer ytd-channel-name yt-formatted-string") |> html_text(trim = TRUE)
  channels <- channels_raw[seq(1, length(channels_raw), by = 2)]
  metadata <- html |> html_elements("ytd-video-renderer #metadata-line span") |> html_text(trim = TRUE)

  tibble(
    query   = query,
    rank    = seq_along(titles),
    title   = titles,
    channel = channels,
    views   = metadata[seq(1, length(metadata), by = 2)],
    url     = html |>
                html_elements("ytd-video-renderer #video-title") |>
                html_attr("href") |>
                (\(x) paste0("https://www.youtube.com", x))()
  )
}

queries <- c("Klimawandel", "climate change")

all_results <- map_dfr(queries, \(q) {
  message("Searching: ", q)
  Sys.sleep(1)
  collect_search_results(q, n_scrolls = 3)
})

all_results |>
  count(query, channel, sort = TRUE) |>
  pivot_wider(names_from = query, values_from = n, values_fill = 0)
```

### Step 6: Close the session {#p4-step6}

```{r}
#| label: close
close_session(session)
```

---

## Comparing approaches: API vs. browser automation

| | Undocumented API | Browser Automation |
|---|---|---|
| **Speed** | Fast (milliseconds per request) | Slow (seconds per page) |
| **Reliability** | High (structured data) | Medium (DOM can change) |
| **Setup effort** | Low–Medium | Medium |
| **Works on dynamic sites** | Sometimes | [60% of the time it works all the time](https://www.youtube.com/shorts/UDSH94tdGVU) |
| **Best for** | Search results, large datasets | Forms, login-gated, dynamic pages |

**Rule of thumb**: Always look for an API first. Only use browser automation when you have to.

## Exercises {#exercises-p4}

**Exercise 1 (easy) — Extract the intro paragraph from a Wikipedia page**

Use `selenider` to navigate to the Wikipedia page for the Hertie School and extract the text of the first paragraph of the article body. This is a simple navigation + extraction task with no dynamic content — but it shows the basic `open_url` → `get_page_source()` → `rvest` workflow.

::: {.callout-note collapse="true"}
## Solution

```r
session <- selenider_session("chromote",
  options = chromote_options(headless = FALSE, width = 1280, height = 900))

open_url("https://en.wikipedia.org/wiki/Hertie_School")
Sys.sleep(2)

html <- get_page_source()

paras <- html |>
  html_elements(".mw-parser-output > p") |>
  html_text(trim = TRUE)

paras[nzchar(paras)][1]
# "The Hertie School is a private research university in Berlin, Germany..."

close_session(session)
```

Note: for static pages like Wikipedia, this is illustrative only — `read_html("https://en.wikipedia.org/wiki/Hertie_School")` would give you the exact same result without needing a browser. Browser automation adds value only when content is injected by JavaScript after load.
:::

---

**Exercise 2 (intermediate) — Compare YouTube results for two queries**

Use `collect_search_results()` to collect the top results for two related queries — for example, "nuclear energy" and "Atomkraft". Which channels appear in both result sets? Are there channels that appear only in one language's results?

::: {.callout-note collapse="true"}
## Solution

```r
results <- map_dfr(
  c("nuclear energy", "Atomkraft"),
  \(q) { Sys.sleep(2); collect_search_results(q, n_scrolls = 2) }
)

# Channels in both
shared <- results |>
  group_by(channel) |>
  summarise(queries = n_distinct(query), .groups = "drop") |>
  filter(queries == 2)

shared
# Likely: DW, Euronews, major science channels appear in both
# language-specific channels (ARD, ZDF, BR) appear only in "Atomkraft"
# English-only YouTubers appear only in "nuclear energy"

# Rank of shared channels per query
results |>
  filter(channel %in% shared$channel) |>
  select(query, rank, channel) |>
  arrange(query, rank)
```

The comparison reveals language-segmented information ecosystems: some channels (e.g. DW, which publishes in both German and English) bridge both, while most channels are only surfaced for one language query. This is the basic methodology behind DSA-style algorithmic audits — systematic comparison of what a platform recommends for equivalent queries across different languages or regions.
:::

---

## ✋ Your turn {#your-turn-p4}

**Option A — YouTube search audit (recommended):**  
Choose a policy topic you care about — migration, nuclear energy, vaccine hesitancy, a recent national election. Pick 2–3 search queries for the same topic (e.g., the same term in English, German, and French). Use the `collect_search_results()` function to gather the top ~40 results for each query. Which channels appear consistently across languages? Are the same videos surfaced? Write 2–3 sentences describing what this might reveal about YouTube's recommendation logic.

**Option B — German Federal Gazette (Bundesanzeiger):**  
[bundesanzeiger.de](https://www.bundesanzeiger.de) publishes mandatory corporate filings through a JavaScript-rendered interface. Use `selenider` to fill the search form, wait for results to appear, and collect company names, filing types, and publication dates. (Note: the site uses a CAPTCHA on some search paths — try company name search first.)

For either option:
1. Open the Network tab first and check whether a hidden API exists — it is always faster to use an API if one is available.
2. Save your results to a CSV.
3. Write 2–3 sentences on what story you could tell with this data.

---

# Putting it together: a mini-investigation

Here is a realistic workflow that combines both techniques: use the Wikipedia REST API to collect a set of articles programmatically, then use browser automation to extract the full rendered content from their detail pages — which includes infobox data and section structure that the search API doesn't return.

::: {.callout-note}
## Browser automation here is illustrative only

Wikipedia article pages are **static HTML** — they do not require JavaScript to render their content. Using browser automation to retrieve them is therefore unnecessary overhead. In a real project you would download each page directly with `httr2` and parse it with `rvest::read_html()`, iterating over URLs in a loop or with `purrr::map()`. See the [session on scraping multiple pages](https://raw.githack.com/intro-to-data-science-25/labs/main/session-05-web-scraping/5-web-scraping.html) from the IDS lab materials for a clean template. Browser automation is retained here solely to show how the two techniques can be combined in a single workflow.
:::

```{r}
#| label: mini-investigation

library(httr2)
library(selenider)
library(rvest)
library(tidyverse)

# Step 1: Use the Wikipedia search API to find articles about
# public policy schools in Europe
search_results <-
  request("https://en.wikipedia.org/w/rest.php/v1/search/page") |>
  req_url_query(q = "public policy school Europe", limit = 20) |>
  req_headers(`User-Agent`      = "DataJournalismCourse/1.0",
              `Accept-Encoding` = "gzip, deflate") |>
  req_perform() |>
  resp_body_json(simplifyVector = TRUE) |>
  pluck("pages") |>
  select(key, title, description, excerpt)

# Step 2: For the top results, use browser automation to retrieve
# full infobox data from each article page.
# Note: Wikipedia article pages are *static* HTML — browser automation is not
# needed here and is used only to illustrate the combined workflow. In practice,
# the better approach is to download the HTML directly with httr2 or read_html()
# and parse with rvest. See the "Scraping multiple pages" example in the web
# scraping lab:
# https://raw.githack.com/intro-to-data-science-25/labs/main/session-05-web-scraping/5-web-scraping.html

session <- selenider_session("chromote", options = chromote_options(headless = TRUE, width = 1280, height = 900))

get_infobox <- function(article_key) {
  open_url(paste0("https://en.wikipedia.org/wiki/", article_key))
  Sys.sleep(2)

  html <- get_page_source()

  labels <- html |>
    html_elements(".infobox th") |>
    html_text(trim = TRUE)

  values <- html |>
    html_elements(".infobox td") |>
    html_text(trim = TRUE)

  if (length(labels) == 0) return(NULL)

  tibble(
    article = article_key,
    label   = labels[seq_len(min(length(labels), length(values)))],
    value   = values[seq_len(min(length(labels), length(values)))]
  )
}

infoboxes <- map(
  head(search_results$key, 10),
  \(key) {
    message("Fetching: ", key)
    Sys.sleep(1)
    get_infobox(key)
  }
) |>
  compact() |>
  list_rbind()

close_session(session)

# Step 3: Analyse — what do the infoboxes tell us?
infoboxes |>
  filter(str_detect(label, regex("established|founded|endow", ignore_case = TRUE))) |>
  pivot_wider(names_from = label, values_from = value) |>
  print(n = 20)
```

---

# Resources

- Leon Yin, *Inspect Element* — [inspectelement.org](https://inspectelement.org)  
  The best practical guide to this workflow, in Python. The undocumented API section in this tutorial is mainly inspired by this guide.

- `httr2` documentation — [httr2.r-lib.org](https://httr2.r-lib.org/)  
  The modern R HTTP client. Read especially the articles on [requests](https://httr2.r-lib.org/articles/httr2.html) and [error handling](https://httr2.r-lib.org/reference/req_retry.html).

- `selenider` documentation — [ashbythorpe.github.io/selenider](https://ashbythorpe.github.io/selenider/)  
  R browser automation with a clean tidyverse-style API.

- [curlconverter.com](https://curlconverter.com/) — converts cURL commands to R (`httr2`), Python, and many other languages.

- The Markup, *[How we investigated Google](https://themarkup.org/google-the-giant/2020/07/28/how-we-analyzed-google-search-results-web-assay-parsing-tool)* — a detailed methodology on using browser automation in a real investigation.

---

::: {.callout-note collapse="true"}
## Session info

```{r}
#| eval: true
sessionInfo()
```
:::
